---
title: "Probabilistic Learning"
output: html_document
---

# Naive Bayes

**Bayesian methods** describe the probability of events, and how probabilities should be revised in the light of additional information.

Classifiers based on Bayesian methods utilize training data to calculate an observed probability of each outcome based on the evidence provided by feature values.

Bayesian probability theory is rooted in the idea that the estimated likelihood of an event, or a potential outcome, should be based on the evidence at hand across multiple trials, or opportunities for the event to occur.

## Bayes Theorem

For independent events $A$ and $B$, the probability of both happening can be expressed as $P(A \cap B) = P(A) * P(B)$.

The relationships between dependent events can be described using **Bayes theorem**:

$$
P(A|B) = \frac{P(A ∩ B)}{P(B)}
$$

The notation $P(A|B)$ is read as the probability of event $A$, given that event $B$ occurred. This is known as **conditional probability**.

## The Naive Bayes algorithm

| **Strengths**                                                                                          | **Weaknesses**                                                                     |
|--------------------------------------------------------------------------------------------------------|------------------------------------------------------------------------------------|
| Simple, fast, and very effective                                                                       | Relies on an often-faulty assumption of equally important and independent features |
| Does well with noisy and missing data                                                                  | Not ideal for datasets with many numeric features                                  |
| Requires relatively few examples for training, but also works well with very large numbers of examples | Estimated probabilities are less reliable than the predicted classes               |
| Easy to obtain the estimated probability for a prediction                                              |                                                                                    |

If $A$ depends on a number of events we can replace $B$ with the intersection of these events, for example:

$$
P(A|B_1 \cap B_2 \cap B_3) = \frac{P(B_1 \cap B_2 \cap B_3|A) P(A)}{P(B_1 \cap B_2 \cap B_3)}
$$

Then if we assume the events are independent, this can simplify to:

$$
P(A|B_1 \cap B_2 \cap B_3) = \frac{P(B_1|A) P(B_2|A) P(B_3|A) P(A)}{P(B_1 \cap B_2 \cap B_3)}
$$

NOTE: if an event is not recorded in the training data the probability will be set to $0$. Since the probabilities are multiplied this gives an unrealistic answer. The approach to correct this is to add a **Laplace estimator** to the counts in the frequency table, usually $1$. Events not seen in the data then have a small probability of occurrence, rather than no chance.

NOTE: The naive Bayes algorithm only works for categorical features. For numeric features, the usual approach is to split it into **bins**.

## Example

```{r}
library(tidyverse)

sms_raw <- read_csv("sms_spam.csv") |> 
  mutate(type = factor(type))

glimpse(sms_raw)

table(sms_raw$type)
```

### Cleaning Data

```{r}
library(tidytext)
library(SnowballC)

sms_clean <- sms_raw |> 
  rowid_to_column("text_id") |> 
  unnest_tokens(word, text) |> 
  filter(!str_detect(word, "[:digit:]")) |> 
  anti_join(get_stopwords(), by = "word") |> 
  mutate(word = wordStem(word)) |> 
  filter(word != "")

sms_freq_words <- sms_clean |> 
  count(word, sort = TRUE) |> 
  filter(n > 5) |> 
  pull(word)

sms_clean <- sms_clean |> 
  filter(word %in% sms_freq_words)

```

### Analysing Data

```{r}
sms_spam_count <- sms_clean |> 
  filter(type == "spam") |> 
  count(word, sort = TRUE)

sms_spam_count
```

```{r}
sms_ham_count <- sms_clean |> 
  filter(type == "ham") |> 
  count(word, sort = TRUE)

sms_ham_count
```

```{r}
library(ggplot2)
library(ggwordcloud)

sms_spam_count |> 
  slice(1:50) |> 
  ggplot() + 
  geom_text_wordcloud(aes(label = word, size = n)) +
  scale_size_area(max_size = 20)
```

```{r}
sms_ham_count |> 
  slice(1:50) |> 
  ggplot() + 
  geom_text_wordcloud(aes(label = word, size = n)) +
  scale_size_area(max_size = 20)
```

### Preparing Data

```{r}
sms_wide <- sms_clean |> 
  mutate(occurance = "Yes") |> 
  group_by(text_id, word) |> 
  summarise(.type = first(type), occurance = first(occurance), .groups = "drop") |> 
  pivot_wider(
    names_from  = word, 
    values_from = occurance, 
    values_fill = "No"
  ) |> 
  select(-text_id)
  
sms_train <- slice(sms_wide, 1:4169)
sms_test  <- slice(sms_wide, -(1:4169))

```

### Applying Model

```{r}
library(tidymodels)
library(discrim)
library(gmodels)
tidymodels_prefer()

sms_bayes_fit <- 
  naive_Bayes(Laplace = 1) |> 
  set_mode("classification") |> 
  set_engine("naivebayes") |> 
  fit(.type ~ ., data = sms_train)

sms_pred <- sms_bayes_fit |> 
  predict(sms_test) |> 
  bind_cols(sms_test)

CrossTable(
  x = sms_pred$.pred_class,
  y = sms_pred$.type, 
  prop.chisq = FALSE, 
  prop.t = FALSE,
  dnn = c('predicted', 'actual')
)
```
